%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{User-written functions of the CG module}

Due to the relative simplicity of the cut generator, there are no wrapper
functions implemented for CG. Consequently, there are no default
options and no post-processing.

%begin{latexonly}
\bd
%end{latexonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% user_receive_cg_data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\functiondef{user\_receive\_cg\_data}
\label{user_receive_cg_data}
\begin{verbatim}
int user_receive_cg_data (void **user)
\end{verbatim}

\bd

\describe

The user has to receive here all problem-specific information that is
known to the master and will be needed for computation in the CG
module later on. The same data must be received here that was sent in
the \hyperref{{\tt user\_send\_cg\_data()}}{{\tt
user\_send\_cg\_data()} (see Section }{)}{user_send_cg_data} function
in the master module. The user has to allocate space for all the data
structures, including {\tt user} itself. Note that some or all of this
may be done in the function {\tt 
\htmlref{user\_send\_cg\_data()}{user_send_cg_data}} if the Tree Manager, LP,
and CG are all compiled together. See that function for more
information.

\args

\bt{llp{250pt}}
{\tt void **user} & INOUT & Pointer to the user-defined data structure. \\
\et

\returns

\bt{lp{300pt}}
{\tt ERROR} & Error. CG exits. \\
{\tt USER\_NO\_PP} & The user received the data properly. \\
\et

\item[Invoked from:] {\tt cg\_initialize()} at process start.

\ed

\vspace{1ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% user_unpack_lp_solution_cg
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\functiondef{user\_receive\_lp\_solution\_cg}
\begin{verbatim}
int user_receive_lp_solution_cg(void *user)
\end{verbatim}

\bd

\describe

This function is invoked only if in the {\tt
\htmlref{user\_send\_lp\_solution()}{user_send_lp_solution}} function
of the LP module the user opted for packing the current LP solution
himself. Here he must unpack the very same data he packed there.

\args

\bt{llp{250pt}}
{\tt void *user} & IN & Pointer to the user-defined data structure. \\
\et

\item[Invoked from:] Whenever an LP solution is received.

\returns

\bt{lp{300pt}}
{\tt ERROR} & Error. This LP solution is not processed. \\
{\tt USER\_NO\_PP} & The user received the LP solution. \\
\et

\item[Note:] \hfill

\BB\ automatically unpacks the level, index and iteration number
corresponding to the current LP solution within the current search tree node
as well as the objective value and upper bound.

\ed

\vspace{1ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% user_free_cg
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\functiondef{user\_free\_cg}
\begin{verbatim}
int user_free_cg(void **user)
\end{verbatim}

\bd

\describe

The user has to free all the data structures within {\tt user}, and also free
{\tt user} itself. The user can use the built-in macro {\tt FREE} that checks
the existence of a pointer before freeing it. 

\args

\bt{llp{280pt}}
{\tt void **user} & INOUT & Pointer to the user-defined data structure
(should be {\tt NULL} on exit from this function). \\
\et

\returns

\bt{lp{300pt}}
{\tt ERROR} & Ignored. \\
{\tt USER\_NO\_PP} & The user freed all data structures. \\
\et

\item[Invoked from:] {\tt cg\_close()} at module shutdown. 

\ed

\vspace{1ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% user_find_cuts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\functiondef{user\_find\_cuts}
\begin{verbatim}
int user_find_cuts(void *user, int varnum, int iter_num, int level,
		    int index, double objval, int *indices, double *values,
		    double ub, double lpetol, int *cutnum)
\end{verbatim}

\bd

\describe

The user can generate cuts based on the current LP solution stored in
{\tt soln}. Cuts found need to be sent back to the LP by calling the
{\tt cg\_send\_cut(\htmlref{cut\_data}{cut_data} *new\_cut)} function.
The argument of this function is a pointer to the cut to be sent. See
Section
\ref{user-written-lp} for a description of this data structure. If the
user wants the cut to be added to the cut pool in case it proves to be
effective in the LP, then {\tt new\_cut->name} should be set to {\tt
CUT\_\_SEND\_TO\_CP}. Otherwise, it should be set to {\tt
CUT\_\_DO\_NOT\_SEND\_TO\_CP}.\\
\\
The only output of this function is the number of cuts generated and this
value is returned in the last argument.

\args

\bt{llp{280pt}}
{\tt void *user} & IN & Pointer to the user-defined data structure.
\\
{\tt int iter\_num} & IN & The iteration number of the current LP solution. \\
{\tt int level} & IN & The level in the tree on which the current LP
solution was generated. \\
{\tt index} & IN & The index of the node in which LP solution was generated.
\\
{\tt objval} & IN & The objective function value of the current LP solution.
\\
{\tt int varnum} & IN & The number of nonzeros in the current LP solution. \\
{\tt indices} & IN & The column indices of the nonzero variables in the current
LP solution. \\
{\tt values} & IN & The values of the nonzero variables listed in 
{\tt indices}.
\\
{\tt double ub} & IN & The current global upper bound. \\
{\tt double lpetol} & IN & The current error tolerance in the LP. \\
{\tt int *cutnum} & OUT & Pointer to the number of cuts generated
and sent to the LP. \\
\et

\returns

\bt{lp{300pt}}
{\tt ERROR} & Ignored. \\
{\tt USER\_NO\_PP} & The user function exited properly. \\
\et

\item[Invoked from:] Whenever an LP solution is received.

\ed

\vspace{1ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% user_check_validity_of_cut
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{user_check_validity_of_cut}
\functiondef{user\_check\_validity\_of\_cut}
\begin{verbatim}
int user_check_validity_of_cut(void *user, cut_data *new_cut)
\end{verbatim}

\bd

\describe

This function is provided as a debugging tool. Every cut that is to be
sent to the LP solver is first passed to this function where the user
can independently verify that the cut is valid by testing it against a
known feasible solution (usually an optimal one). This is useful for
determining why a particular known feasible (optimal) solution was
never found. Usually, this is due to an invalid cut being added. See
Section \ref{debugging} for more on this feature.

\args

\bt{llp{280pt}}
{\tt void *user} & IN & Pointer to the user-defined data structure. \\
{\tt cut\_data *new\_cut} & IN & Pointer to the cut that must be
checked. \\
\et

\returns

\bt{lp{300pt}}
{\tt ERROR} & Ignored. \\
{\tt USER\_NO\_PP} & The user is done checking the cut. \\
\et

\item[Invoked from:] Whenever a cut is being sent to the LP.

\ed

\vspace{1ex}

%begin{latexonly}
\ed
%end{latexonly}